### 2020-02-04 paper/introduction.tex
revmove definition for SD-TMSR

Various previous works explore starting the
\glspl{MSR} with fissile materials alternative to $^{233}$U. Many such publications have focused on the fast-spectrum \glspl{MSR} \cite{heuer2014towards,ashraf2019modeling,
	ashraf2018nuclear, rykhlevskii_fuel_2019, betzler_impacts_2019,
	fiorina2013investigation}, while few focus on
thermal-spectrum \glspl{MSR} \cite{betzler2016modeling,
	zou2018transition,zou2018preliminary}. Therefore,
the main objective of the present paper is to establish feasible strategies for thorium fuel cycle transition in the \gls{SD-TMSR} with
various initial fissile materials and without any external feed of $^{233}$U. We investigate five different
initial fissile materials: \gls{HALEU}, Pu mixed with \gls{HALEU}, reactor-grade Pu \footnote{Reactor-grade Pu is a mixture of Pu isotopes chemically extracted from PWR SNF with 33 GWd/tHM burnup} \cite{marka1993explosive}, \gls{TRU} from LWR SNF \cite{de2000scenarios}, and $^{233}$U. Two different feed mechanisms were selected:

>>>>

Various previous works explore starting the
\glspl{MSR} with fissile materials alternative to $^{233}$U. Many such publications have focused on the fast-spectrum \glspl{MSR} \cite{heuer2014towards,ashraf2019modeling,
	ashraf2018nuclear, rykhlevskii_fuel_2019, betzler_impacts_2019,
	fiorina2013investigation}, while few focus on
thermal-spectrum \glspl{MSR} \cite{betzler2016modeling,
	zou2018transition,zou2018preliminary}. Therefore,
the main objective of the present paper is to establish feasible strategies for thorium fuel cycle transition in the SD-TMSR with
various initial fissile materials and without any external feed of $^{233}$U. We investigate five different
initial fissile materials: \gls{HALEU}, Pu mixed with \gls{HALEU}, reactor-grade Pu \footnote{Reactor-grade Pu is a mixture of Pu isotopes chemically extracted from PWR SNF with 33 GWd/tHM burnup} \cite{marka1993explosive}, \gls{TRU} from LWR SNF \cite{de2000scenarios}, and $^{233}$U. Two different feed mechanisms were selected:

### 2020-02-04 paper/methods.tex
revmove definition for SD-TMSR

Simulation of liquid-fueled Molten Salt Reactor (MSR) systems requires 
computational software that can support online fuel salt reprocessing and 
refueling \cite{serp2014molten}. In this work, SERPENT-2 version 2.1.31 \cite{leppanen2014serpent} simulates the 
Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) full-core with various initial fuel types. This extension 
of SERPENT accounts for continuous online reprocessing and refueling 
\cite{aufiero2013extended}. The ENDF-VII.0 cross section library was used for all 
calculations in this work. The results demonstrate full-core runs of
$1.25\times 10^7$ neutron histories per depletion step. During the depletion step, the core was maintained critical and total fuel mass was almost constant (dm $\leq$ 0.1\%). To determine the appropriate depletion step size, we conducted a time step refinement study. We started with $\Delta t=30$days and gradually increased the depletion time step until the error in K$_{eff}$ became significant. For $\Delta t=1$year the maximum change in K$_{eff}$ was $\leq$ 0.02\%), therefore we adopted a constant, 365-day-long time step for all simulation in this work.
The full burnup time of 
the SD-TMSR was 60 years with statistical error in $k_{eff}$ equal to $\pm$ 
$12$ $pcm$. 
During the molten salt reactor operation, part of fuel salt flows outside of the reactor core then the fission reaction has not happened in the external loop, however, the decay reactions happen everywhere. We used total fuel salt inventory in the primary loop to define the fuel material in the SERPENT input. Moreover, we accordingly adjusted power density in SERPENT burnup calculations because fission takes place only in the core region.

>>>>

Simulation of liquid-fueled Molten Salt Reactor (MSR) systems requires 
computational software that can support online fuel salt reprocessing and 
refueling \cite{serp2014molten}. In this work, SERPENT-2 version 2.1.31 \cite{leppanen2014serpent} simulates the 
SD-TMSR full-core with various initial fuel types. This extension 
of SERPENT accounts for continuous online reprocessing and refueling 
\cite{aufiero2013extended}. The ENDF-VII.0 cross section library was used for all 
calculations in this work. The results demonstrate full-core runs of
$1.25\times 10^7$ neutron histories per depletion step. During the depletion step, the core was maintained critical and total fuel mass was almost constant (dm $\leq$ 0.1\%). To determine the appropriate depletion step size, we conducted a time step refinement study. We started with $\Delta t=30$days and gradually increased the depletion time step until the error in K$_{eff}$ became significant. For $\Delta t=1$year the maximum change in K$_{eff}$ was $\leq$ 0.02\%), therefore we adopted a constant, 365-day-long time step for all simulation in this work.
The full burnup time of 
the SD-TMSR was 60 years with statistical error in $k_{eff}$ equal to $\pm$ 
$12$ $pcm$. 
During the molten salt reactor operation, part of fuel salt flows outside of the reactor core then the fission reaction has not happened in the external loop, however, the decay reactions happen everywhere. We used total fuel salt inventory in the primary loop to define the fuel material in the SERPENT input. Moreover, we accordingly adjusted power density in SERPENT burnup calculations because fission takes place only in the core region.

### 2020-02-04 paper/results.tex
revmove definition for SD-TMSR

\section{Results and discussion} \label{Results-and-discussion}
The molar fraction of the heavy metal (HM) 
in the initial fuel was kept constant and equal to 12.5 mole\% for all cases (see Table~\ref{tab:table4}). 
Additionally, the initial fissile material fraction was increased for the five fuel 
salt compositions until the \gls{SD-TMSR} reactor was sufficiently critical at 
the Beginning Of Life (BOL).
\subsection{Thorium feed mechanism}
The thorium feed mechanism continuously feeds thorium from the external stockpile and 
$^{233}$U from the \texttt{Pa-decay tank}.
Figure~\ref{fig:keff1} illustrates the effective multiplication factor 
dynamics during reactor operation for the thorium feed mechanism. As shown in 
Figure~\ref{fig:keff1}, the effective multiplication factor ($k_{eff}$) 
decreases sharply during the first 25 \gls{EFPY} of reactor operation for the first 
four cases. $k_{eff}$ decreases as a result of depletion of the initial 
fissile materials and production of poisonous fission products (FPs). The amount of $^{233}$U generated in 
the SD-TMSR is not enough to maintain the reactor criticality and 
counteract parasitic neutron absorption. Thus, the 
reactor becomes subcritical relatively quickly for alternative startup 
compositions, for example, $\approx$ $4$ years in the \gls{TRU} case and $\approx$ $12$ 
years in the reactor-grade Pu case. For U-233 case, 
the continuous feed flow of thorium and $^{233}$U from the \texttt{Pa-decay tank} helps to operate the 
\gls{SD-TMSR} for a full lifetime (60 year) (Figure~\ref{fig:keff1}, U-233 case).

@@

\subsection{Non-thorium feed mechanism}
The non-thorium feed mechanism continuously feeds 
$^{233}$U from the \texttt{Pa-decay tank} and external heavy metals (\gls{HALEU}, Pu mixed with \gls{HALEU}, reactor-grade Pu, \gls{TRU}).
Under the non-thorium feed mechanism, only four different initial fissile materials are studied: \gls{HALEU}, Pu mixed with \gls{HALEU}, reactor-grade Pu, and \gls{TRU}. The continuous feed of $^{233}$U 
without $^{232}$Th will lead to a supercritical reactor, thus the $^{233}$U case 
is excluded from the non-thorium feed mechanism study.
Figure~\ref{fig:keff2} shows the change of the effective multiplication factor 
during 60 \gls{EFPY} of reactor operation for the non-thorium feed mechanism. Both the reactor-grade Pu and TRU cases 
show promising results relative to the other two cases (HALEU and Pu+HALEU) (see Figure~\ref{fig:keff2}). 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{keff2.png}
			\vspace{-0.5in}
	\caption{The change of the effective multiplication factor during 60 
	\gls{EFPY} of reactor operation for non-thorium feed mechanism (confidence 
	interval $\pm\sigma$ is shaded).} 
	\label{fig:keff2}
\end{figure}
For the reactor-grade Pu case, the amount of $^{233}$U generated in the \gls{SD-TMSR}, in addition to the external feed flow of Pu, is sufficient to maintain the reactor criticality and counteract the neutron absorption in the initial non-fissile isotopes and FPs. 
For the \gls{TRU} fuel salt, the amount of 
$^{233}$U and the external feed flow of TRU is barely enough to operate the 
reactor for a long period of time ($\approx$ $40$ years) without any 
external feed of $^{233}$U ($^{233}$U used only from the \texttt{Pa-decay tank}). Nevertheless, $k_{eff}$ decreases with the burnup 
because of the \glspl{MA}\footnote{In the present work, the minor actinides 
(MA) include Np, Am, and Cm.} accumulating in the core as a result of 
continuous TRU feed. As shown in Figure~\ref{fig:keff2}, the HALEU and Pu+HALEU 
fuel are less attractive for the non-thorium feed mechanism. The continuous HALEU 
feed increases the amount of fertile $^{238}$U and consequently, reduces the 
feasibility of such fissile materials. According to the $k_{eff}$ 
results, reactor-grade Pu and TRU are the only alternative fissile materials that  
can be used to startup and maintain operation of the \gls{SD-TMSR}.

@@

Figure~\ref{fig:NetCC} demonstrates the net production of $^{233}$U during 
reactor operation for the $^{233}$U, reactor-grade Pu, and TRU cases, respectively. For the TRU case, the net production of $^{233}$U is almost 
zero. 
Although all produced $^{233}$U is used to refuel the core, the reactor is subcritical after $40$ years of operation (see Figure~\ref{fig:keff2}).
In the $^{233}$U and reactor-grade Pu cases, the net production of 
$^{233}$U increases with burnup and reaches about $1.77$ $t$ and $10$ $t$ at EOL, respectively. Figure~\ref{fig:NetCC} shows that for the $^{233}$U 
case the net production of $^{233}$U during the first 455 days is negative; 
about $175.28$ $kg$ of $^{233}$U must be added during this period. 
As shown in Figure~\ref{fig:NetCC}, for the $^{233}$U case, after 26 years the net production of $^{233}$U reaches 
$1.3$ $t$; this is sufficient to start up another \gls{SD-TMSR} (see Table~\ref{tab:table5}). Similarly, 
one can see that the same amount of $^{233}$U ($1.3$ $t$) can be achieved 
after $\approx$ $4.5$ years if we apply the non-thorium feed mechanism on 
the SD-TMSR that was initially loaded by reactor-grade Pu alternative to 
$^{233}$U. For nonproliferation reasons, $^{233}$U could be diluted with $^{238}$U which makes it impossible to use for nuclear weapons.

@@

\subsection{Neutron spectrum}
Figure~\ref{fig:spectrumFLUX110vC} represents the neutron flux per unit 
lethargy for a full-core SD-TMSR model in the energy range from 10$^{-8}$ to 
10 MeV for the $^{233}$U, reactor-grade Pu, and TRU cases at BOL and EOL. In 
the $^{233}$U case, at the EOL, the neutron spectrum is harder than at BOL due 
to the accumulation of Pu and other strong thermal neutron absorbers in the 
fuel salt. For the reactor-grade Pu and TRU cases, during reactor operation, 
the fissile Pu is depleted and the $^{233}$U becomes the major fissile isotope 
(see Figure~\ref{fig:U233CC}); the neutron spectrum softens and becomes 
similar to the initial thermal spectrum of a $^{233}$U fueled \gls{SD-TMSR}.
 \begin{figure}
 	\centering
 	\includegraphics[width=1.01\textwidth]{spectrumFLUX110vC.png}
 			\vspace{-0.4in}
 	\caption{The neutron flux energy spectrum at BOL (solid lines) and EOL (dashed lines) for 3 different initial 
 		fuel salt compositions (for all cases, the neutron flux
 		confidence intervals $\pm\sigma$ at BOL and EOL are $<1.0139$\% and $<0.6274$\%, respectively).}
 	\label{fig:spectrumFLUX110vC}
\end{figure}

>>>>

\section{Results and discussion} \label{Results-and-discussion}
The molar fraction of the heavy metal (HM) 
in the initial fuel was kept constant and equal to 12.5 mole\% for all cases (see Table~\ref{tab:table4}). 
Additionally, the initial fissile material fraction was increased for the five fuel 
salt compositions until the SD-TMSR reactor was sufficiently critical at 
the Beginning Of Life (BOL).
\subsection{Thorium feed mechanism}
The thorium feed mechanism continuously feeds thorium from the external stockpile and 
$^{233}$U from the \texttt{Pa-decay tank}.
Figure~\ref{fig:keff1} illustrates the effective multiplication factor 
dynamics during reactor operation for the thorium feed mechanism. As shown in 
Figure~\ref{fig:keff1}, the effective multiplication factor ($k_{eff}$) 
decreases sharply during the first 25 \gls{EFPY} of reactor operation for the first 
four cases. $k_{eff}$ decreases as a result of depletion of the initial 
fissile materials and production of poisonous fission products (FPs). The amount of $^{233}$U generated in 
the SD-TMSR is not enough to maintain the reactor criticality and 
counteract parasitic neutron absorption. Thus, the 
reactor becomes subcritical relatively quickly for alternative startup 
compositions, for example, $\approx$ $4$ years in the \gls{TRU} case and $\approx$ $12$ 
years in the reactor-grade Pu case. For U-233 case, 
the continuous feed flow of thorium and $^{233}$U from the \texttt{Pa-decay tank} helps to operate the 
SD-TMSR for a full lifetime (60 year) (Figure~\ref{fig:keff1}, U-233 case).

@@

\subsection{Non-thorium feed mechanism}
The non-thorium feed mechanism continuously feeds 
$^{233}$U from the \texttt{Pa-decay tank} and external heavy metals (\gls{HALEU}, Pu mixed with \gls{HALEU}, reactor-grade Pu, \gls{TRU}).
Under the non-thorium feed mechanism, only four different initial fissile materials are studied: \gls{HALEU}, Pu mixed with \gls{HALEU}, reactor-grade Pu, and \gls{TRU}. The continuous feed of $^{233}$U 
without $^{232}$Th will lead to a supercritical reactor, thus the $^{233}$U case 
is excluded from the non-thorium feed mechanism study.
Figure~\ref{fig:keff2} shows the change of the effective multiplication factor 
during 60 \gls{EFPY} of reactor operation for the non-thorium feed mechanism. Both the reactor-grade Pu and TRU cases 
show promising results relative to the other two cases (HALEU and Pu+HALEU) (see Figure~\ref{fig:keff2}). 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{keff2.png}
			\vspace{-0.5in}
	\caption{The change of the effective multiplication factor during 60 
	\gls{EFPY} of reactor operation for non-thorium feed mechanism (confidence 
	interval $\pm\sigma$ is shaded).} 
	\label{fig:keff2}
\end{figure}
For the reactor-grade Pu case, the amount of $^{233}$U generated in the SD-TMSR, in addition to the external feed flow of Pu, is sufficient to maintain the reactor criticality and counteract the neutron absorption in the initial non-fissile isotopes and FPs. 
For the \gls{TRU} fuel salt, the amount of 
$^{233}$U and the external feed flow of TRU is barely enough to operate the 
reactor for a long period of time ($\approx$ $40$ years) without any 
external feed of $^{233}$U ($^{233}$U used only from the \texttt{Pa-decay tank}). Nevertheless, $k_{eff}$ decreases with the burnup 
because of the \glspl{MA}\footnote{In the present work, the minor actinides 
(MA) include Np, Am, and Cm.} accumulating in the core as a result of 
continuous TRU feed. As shown in Figure~\ref{fig:keff2}, the HALEU and Pu+HALEU 
fuel are less attractive for the non-thorium feed mechanism. The continuous HALEU 
feed increases the amount of fertile $^{238}$U and consequently, reduces the 
feasibility of such fissile materials. According to the $k_{eff}$ 
results, reactor-grade Pu and TRU are the only alternative fissile materials that  
can be used to startup and maintain operation of the SD-TMSR.

@@

Figure~\ref{fig:NetCC} demonstrates the net production of $^{233}$U during 
reactor operation for the $^{233}$U, reactor-grade Pu, and TRU cases, respectively. For the TRU case, the net production of $^{233}$U is almost 
zero. 
Although all produced $^{233}$U is used to refuel the core, the reactor is subcritical after $40$ years of operation (see Figure~\ref{fig:keff2}).
In the $^{233}$U and reactor-grade Pu cases, the net production of 
$^{233}$U increases with burnup and reaches about $1.77$ $t$ and $10$ $t$ at EOL, respectively. Figure~\ref{fig:NetCC} shows that for the $^{233}$U 
case the net production of $^{233}$U during the first 455 days is negative; 
about $175.28$ $kg$ of $^{233}$U must be added during this period. 
As shown in Figure~\ref{fig:NetCC}, for the $^{233}$U case, after 26 years the net production of $^{233}$U reaches 
$1.3$ $t$; this is sufficient to start up another SD-TMSR (see Table~\ref{tab:table5}). Similarly, 
one can see that the same amount of $^{233}$U ($1.3$ $t$) can be achieved 
after $\approx$ $4.5$ years if we apply the non-thorium feed mechanism on 
the SD-TMSR that was initially loaded by reactor-grade Pu alternative to 
$^{233}$U. For nonproliferation reasons, $^{233}$U could be diluted with $^{238}$U which makes it impossible to use for nuclear weapons.

@@

\subsection{Neutron spectrum}
Figure~\ref{fig:spectrumFLUX110vC} represents the neutron flux per unit 
lethargy for a full-core SD-TMSR model in the energy range from 10$^{-8}$ to 
10 MeV for the $^{233}$U, reactor-grade Pu, and TRU cases at BOL and EOL. In 
the $^{233}$U case, at the EOL, the neutron spectrum is harder than at BOL due 
to the accumulation of Pu and other strong thermal neutron absorbers in the 
fuel salt. For the reactor-grade Pu and TRU cases, during reactor operation, 
the fissile Pu is depleted and the $^{233}$U becomes the major fissile isotope 
(see Figure~\ref{fig:U233CC}); the neutron spectrum softens and becomes 
similar to the initial thermal spectrum of a $^{233}$U fueled SD-TMSR.
 \begin{figure}
 	\centering
 	\includegraphics[width=1.01\textwidth]{spectrumFLUX110vC.png}
 			\vspace{-0.4in}
 	\caption{The neutron flux energy spectrum at BOL (solid lines) and EOL (dashed lines) for 3 different initial 
 		fuel salt compositions (for all cases, the neutron flux
 		confidence intervals $\pm\sigma$ at BOL and EOL are $<1.0139$\% and $<0.6274$\%, respectively).}
 	\label{fig:spectrumFLUX110vC}
\end{figure}
